Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-29 paper/icml_drau/impl_details.tex
.

%! TEX root = supplementary.tex
All the details specifying our implementation are hereby given. We first give precise pseudo code
descriptions for both \emph{Continuous Deep Advantage Updating} (Alg.~\ref{alg:cdau}), as well as the variants of DDPG (Alg.~\ref{alg:ddpg})
and DQN (Alg.~\ref{alg:dqn}) used.

@@

For DDPG and DQN, two different settings were experimented with:
\begin{itemize}
	\item One with time discretization scalings, to keep the comparison
		fair. In this setting, the discount factor is still scaled as $\gamma^\deltat$,
		rewards are scaled as $r \deltat$, and learning rates are scaled to obtain parameter
		updates of order $\deltat$. As RMSprop is used for all experiments, this amounts
		to using a learning rate scaling as $\alpha^Q = \tilde{\alpha}^Q \deltat$,
		$\alpha^\pi = \tilde{\alpha}^\pi \deltat$.
	\item One without discretization scalings. In that case, only the discount factor is scaled
		as $\gamma^\deltat$, to prevent unfair shortsightedness. All other
		parameters are set with a reference $\deltat_0 = 1e-2$. For instance,
		for all $\deltat$'s, the reward perceived is $r * \deltat_0$, and
		similarily for learning rates, $\alpha^Q = \tilde{\alpha}^Q
		\deltat_0$, $\alpha^\pi = \tilde{\alpha}^Q \deltat_0$. These scalings
		don't depend on the discretization, but perform decently at least for
		the highest discretization.
\end{itemize}
\subsection{Global hyperparameters}
The following hyperparameters are maintained constant throughout all our experiments,
\begin{itemize}
	\item All networks used are of the form
		\begin{verbatim}
		Sequential(
		    Linear(nb_inputs, 256),
		    LayerNorm(256),
		    ReLU(),
		    Linear(256, 256),
		    LayerNorm(256),
		    ReLU(),
		    Linear(256, nb_outputs)
		).
		\end{verbatim}
	Policy networks have an additional $\tanh$ layer to constraint action range. On certain
	environments, network inputs are normalized by applying a mean-std normalization, with
	mean and standard deviations computed on each individual input features, on all previously
	encountered samples.
	\item $\mathcal{D}$ is a cyclic buffer of size $1000000$.
	\item $\textbf{nb\_steps}$ is set to $10$, and $256$ environments are run in parallel to
		accelerate the training procedure, totalling $2560$ environment interactions between
		learning steps.
	\item $\textbf{nb\_learn}$ is set to $50$.
	\item The physical $\gamma$ is set to $0.8$. It is always scaled as $\gamma^\deltat$ (even for
		unscaled DQN and DDPG).
	\item $N$, the batch size is set to $256$.
	\item RMSprop is used as an optimizer without momentum, and with
		$\alpha=1 - \deltat$ (or $1 - \deltat_0$ for unscaled DDPG and
		DQN).
	\item Exploration is always performed as described in the main text. The OU process used as
		parameters $\kappa = 7.5$, $\sigma = 1.5$.
	\item Unless otherwise stated, $\alpha_1 \deq \tilde{\alpha}^Q = \alpha^V = \alpha^A = 0.1$, $\alpha_2 \deq \tilde{\alpha}^\pi =
		\alpha^\pi = 0.03$.
	\item $\tau = 0.9$
\end{itemize}
\subsection{Environment dependent hyperparameters}
We hereby list the hyperparameters used for each environment. Continuous actions environments are marked with a
(C), discrete actions environments with a (D).
\begin{itemize}
	\item {\bf Ant (C)}: State normalization is used. Discretization range: $[0.05, 0.02, 0.01, 0.005, 0.002]$.
	\item {\bf Cheetah (C)}: State normalization is used. Discretization range: $[0.05, 0.02, 0.01, 0.005, 0.002]$
	\item {\bf Bipedal Walker (C)}: State normalization is used, $\alpha_2 = 0.02$. Discretization range: $[0.01, 0.005, 0.002, 0.001]$.
	\item {\bf Cartpole (D)}: $\alpha_2 = 0.02$, $\tau = 0$. Discretization range: $[0.01, 0.005, 0.002, 0.001, 0.0005]$.
	\item {\bf Pendulum (C)}: $\alpha_2 = 0.02$, $\tau = 0$. Discretization range: $[0.01, 0.005, 0.002, 0.001, 0.0005]$.

>>>>

%! TEX root = supplementary.tex
All the details specifying our implementation are given in this section. We
first give precise pseudo code descriptions for both \emph{Continuous Deep
Advantage Updating} (Alg.~\ref{alg:cdau}), as well as the variants of DDPG
(Alg.~\ref{alg:ddpg}) and DQN (Alg.~\ref{alg:dqn}) used.

@@

For DDPG and DQN, two different settings were experimented with:
\begin{itemize}
	\item One with time discretization scalings, to keep the comparison
		fair. In this setting, the discount factor is still scaled as $\gamma^\deltat$,
		rewards are scaled as $r \deltat$, and learning rates are scaled to obtain parameter
		updates of order $\deltat$. As RMSprop is used for all experiments, this amounts
		to using a learning rate scaling as $\alpha^Q = \tilde{\alpha}^Q \deltat$,
		$\alpha^\pi = \tilde{\alpha}^\pi \deltat$.
	\item One without discretization scalings. In that case, only the discount factor is scaled
		as $\gamma^\deltat$, to prevent unfair shortsightedness. All other
		parameters are set with a reference $\deltat_0 = 1e-2$. For instance,
		for all $\deltat$'s, the reward perceived is $r * \deltat_0$, and
		similarily for learning rates, $\alpha^Q = \tilde{\alpha}^Q
		\deltat_0$, $\alpha^\pi = \tilde{\alpha}^Q \deltat_0$. These scalings
		don't depend on the discretization, but perform decently at least for
		the highest discretization.
\end{itemize}
\subsection{Global hyperparameters}
The following hyperparameters are maintained constant throughout all our experiments,
\begin{itemize}
	\item All networks used are of the form
		\begin{verbatim}
		Sequential(
		    Linear(nb_inputs, 256),
		    LayerNorm(256),
		    ReLU(),
		    Linear(256, 256),
		    LayerNorm(256),
		    ReLU(),
		    Linear(256, nb_outputs)
		).
		\end{verbatim}
	Policy networks have an additional $\tanh$ layer to constraint action range. On certain
	environments, network inputs are normalized by applying a mean-std normalization, with
	mean and standard deviations computed on each individual input features, on all previously
	encountered samples.
	\item $\mathcal{D}$ is a cyclic buffer of size $1000000$.
	\item $\textbf{nb\_steps}$ is set to $10$, and $256$ environments are run in parallel to
		accelerate the training procedure, totalling $2560$ environment interactions between
		learning steps.
	\item $\textbf{nb\_learn}$ is set to $50$.
	\item The physical $\gamma$ is set to $0.8$. It is always scaled as $\gamma^\deltat$ (even for
		unscaled DQN and DDPG).
	\item $N$, the batch size is set to $256$.
	\item RMSprop is used as an optimizer without momentum, and with
		$\alpha=1 - \deltat$ (or $1 - \deltat_0$ for unscaled DDPG and
		DQN).
	\item Exploration is always performed as described in the main text. The OU process used as
		parameters $\kappa = 7.5$, $\sigma = 1.5$.
	\item Unless otherwise stated, $\alpha_1 \deq \tilde{\alpha}^Q = \alpha^V = \alpha^A = 0.1$, $\alpha_2 \deq \tilde{\alpha}^\pi =
		\alpha^\pi = 0.03$.
	\item $\tau = 0.9$
\end{itemize}
\subsection{Environment dependent hyperparameters}
We hereby list the hyperparameters used for each environment. Continuous actions environments are marked with a
(C), discrete actions environments with a (D).
\begin{itemize}
	\item {\bf Ant (C)}: State normalization is used. Discretization range: $[0.05, 0.02, 0.01, 0.005, 0.002]$.
	\item {\bf Cheetah (C)}: State normalization is used. Discretization range: $[0.05, 0.02, 0.01, 0.005, 0.002]$
	\item {\bf Bipedal Walker (C)\footnote{
				The reward for Bipedal Walker is modified not to scale with $\deltat$. This does not introduce any change for the default setup.
		}}: State normalization is used, $\alpha_2 = 0.02$. Discretization range: $[0.01, 0.005, 0.002, 0.001]$.
	\item {\bf Cartpole (D)}: $\alpha_2 = 0.02$, $\tau = 0$. Discretization range: $[0.01, 0.005, 0.002, 0.001, 0.0005]$.
	\item {\bf Pendulum (C)}: $\alpha_2 = 0.02$, $\tau = 0$. Discretization range: $[0.01, 0.005, 0.002, 0.001, 0.0005]$.

### 2019-01-29 paper/icml_drau/pendulum_figs_suppl.tex

%! TEX root = supplementary.tex
\newcommand{\iw}{1cm}
\newcommand{\cw}{\textwidth}
\begin{figure*}[h]
	\includegraphics[width=\cw]{figs_data/pendulum_unscaled/pendulum_fig_act_unscaled.pdf}
	\caption{Policies obtained by DDPG (unscaled version) and AU at different instants in physical time of training on the pendulum swing-up environment. Each image represents the policy learnt by the policy network, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the closer to $1$ the action, the darker, the closer to $-1$.}
        \label{fig:pend}
\end{figure*}
\begin{figure*}[h]
	\includegraphics[width=\cw]{figs_data/pendulum_scaled/pendulum_fig_act_scaled.pdf}
	\caption{Policies obtained by DDPG (scaled version) and AU at different instants in physical time of training on the pendulum swing-up environment. Each image represents the policy learnt by the policy network, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the closer to $1$ the action, the darker, the closer to $-1$.}
        \label{fig:pend}
\end{figure*}
\begin{figure*}[h]
	\includegraphics[width=\cw]{figs_data/pendulum_scaled/pendulum_fig_val_scaled.pdf}
	\caption{Value functions obtained by DDPG (scaled version) and AU at different instants in physical time of training on the pendulum swing-up environment. Each image represents the policy learnt by the policy network, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the closer to $1$ the action, the darker, the closer to $-1$.}
        \label{fig:pend}
\end{figure*}

>>>>

%! TEX root = supplementary.tex
\newcommand{\iw}{1cm}
\newcommand{\cw}{\textwidth}
\begin{figure*}[h]
	\includegraphics[width=\cw]{figs_data/pendulum_unscaled/pendulum_fig_act_unscaled.pdf}
	\caption{Policies obtained by DDPG (unscaled version) and AU at different instants in physical time of training on the pendulum swing-up environment. Each image represents the policy learnt by the policy network, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the closer to $1$ the action, the darker, the closer to $-1$.}
	\label{fig:pend1}
\end{figure*}
\begin{figure*}[h]
	\includegraphics[width=\cw]{figs_data/pendulum_scaled/pendulum_fig_act_scaled.pdf}
	\caption{Policies obtained by DDPG (scaled version) and AU at different instants in physical time of training on the pendulum swing-up environment. Each image represents the policy learnt by the policy network, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the closer to $1$ the action, the darker, the closer to $-1$.}
	\label{fig:pend2}
\end{figure*}
\begin{figure*}[h]
	\includegraphics[width=\cw]{figs_data/pendulum_scaled/pendulum_fig_val_scaled.pdf}
	\caption{Value functions obtained by DDPG (scaled version) and AU at different instants in physical time of training on the pendulum swing-up environment. Each image represents the value function learnt, with $x$-axis representing angle, and $y$-axis angular velocity. The lighter the pixel, the higher the value.}
	\label{fig:pend3}
\end{figure*}
\begin{figure*}[h]
	\centering
	\includegraphics[width=\cw]{figs_data/fig_unscaled/full_results_unscaled.pdf}
	\label{fig:full_results}
	\caption{Learning curves for DAU and DDPG (scaled) on classic control benchmarks for  various time discretization $\deltat$: Scaled return as a function of the physical time spent in the environment.}
\end{figure*}

### 2019-01-29 paper/icml_drau/proof_conv_adv.tex

%! TEX root = supplementary.tex
To make things more precise, we denote $\tilde{s}_\deltat^t(s_0)$ the
evaluation at instant $t$ of the solution of $d\tilde{s}_t/dt = F(\tilde{s}_t,
\pi(s_0))$ with starting point $s_0$.

Bellman equation on $Q^\pi_\deltat$ yields
\begin{align}
	Q^\pi_\deltat(s, a) &= r(s, a) \deltat + \gamma^\deltat V^\pi_\deltat(\tilde{s}_\deltat^\deltat(s)).
\end{align}

For all $s$, a first order taylor expansion yields
\begin{equation}
	\tilde{s}_\deltat^\deltat(s) = s + F(s, a)\deltat + \bigO(\deltat^2)
\end{equation}
thus, by uniform boundedness of the hessian of $V^\pi_\deltat$,\footnote{
	Without boundedness of the hessian, we cannot write the second order taylor expansion of
	$V^\pi_\deltat(\tilde{s}_\deltat^\deltat(s))$ in term of $\deltat$.
}
\begin{equation}
	Q^\pi_\deltat(s, a) = r(s, a) \deltat + (1 + \ln(\gamma) \deltat + \bigO(\deltat^2)) (V^\pi_\deltat(s) + \deltat \partial_s V^\pi_\deltat(s) F(s, a) + \bigO(\deltat^2).
\end{equation}
Now, this yields
\begin{equation}
	A^\pi_\deltat(s, a) = r(s, a) + \ln(\gamma) V^\pi_\deltat(s) + \partial_s V^\pi_\deltat(s) F(s, a) + \bigO(\deltat),
\end{equation}
and using the convergence of $V^\pi_\deltat(s)$ to $V^\pi(s)$ (Thm.~\ref{th:conv-value}) and $\partial_s V^\pi_\deltat(s)$ to $\partial_s V^\pi(s)$ (hypothesis) yields the result with 
\begin{equation}
	A^\pi(s, a) = r(s, a) + \ln(\gamma) V^\pi(s) + \partial_s V^\pi(s) F(s, a).
	\label{eq:adv_function}
\end{equation}

>>>>

%! TEX root = supplementary.tex
Denote $\tilde{s}_\deltat^t(s_0)$ the
evaluation at instant $t$ of the solution of $d\tilde{s}_t/dt = F(\tilde{s}_t,
\pi(s_0))$ with starting point $s_0$.

The Bellman equation on $Q^\pi_\deltat$ yields
\begin{align}
	Q^\pi_\deltat(s, a) &= r(s, a) \deltat + \gamma^\deltat V^\pi_\deltat(\tilde{s}_\deltat^\deltat(s)).
\end{align}

For all $s$, a first-order Taylor expansion yields
\begin{equation}
	\tilde{s}_\deltat^\deltat(s) = s + F(s, a)\deltat + \bigO(\deltat^2)
\end{equation}
where the constant in $\bigO()$ is uniformly bounded thanks to the
assumptions on the Hessian.
Thus, by uniform boundedness of the Hessian of $V^\pi_\deltat$,\footnote{
	Without boundedness of the Hessian, we cannot write the second order Taylor expansion of
	$V^\pi_\deltat(\tilde{s}_\deltat^\deltat(s))$ in term of $\deltat$.
}
\begin{equation}
	Q^\pi_\deltat(s, a) = r(s, a) \deltat + (1 + \ln(\gamma) \deltat + \bigO(\deltat^2)) (V^\pi_\deltat(s) + \deltat \partial_s V^\pi_\deltat(s) F(s, a) + \bigO(\deltat^2).
\end{equation}
Now, this yields
\begin{equation}
	A^\pi_\deltat(s, a) = r(s, a) + \ln(\gamma) V^\pi_\deltat(s) + \partial_s V^\pi_\deltat(s) F(s, a) + \bigO(\deltat),
\end{equation}
and using the convergence of $V^\pi_\deltat(s)$ to $V^\pi(s)$ (Thm.~\ref{th:conv-value}) and $\partial_s V^\pi_\deltat(s)$ to $\partial_s V^\pi(s)$ (hypothesis) yields the result with 
\begin{equation}
	A^\pi(s, a) = r(s, a) + \ln(\gamma) V^\pi(s) + \partial_s V^\pi(s) F(s, a).
	\label{eq:adv_function}
\end{equation}

### 2019-01-29 paper/icml_drau/proof_conv_params.tex

%! TEX root = supplementary.tex
Let $(s_t, a_t)_{t\geq 0}$ be the trajectory on which parameters are
learnt. To simplify notations, define
\begin{equation}
	A_\psi(s, a) = \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s)).
\end{equation}
Define $F$ as
\begin{align}
	F^\theta(\theta, \psi, s, a) &= \alpha^V (r(s, a) + \ln(\gamma) V_\theta(s) + \partial_s V_\theta(s) F(s, a) - A_\psi(s, a))\partial_\theta V_\theta(s)\\
	F^\psi(\theta, \psi, s, a) &= \alpha^A (r(s, a) + \ln(\gamma) V_\theta(s) + \partial_s V_\theta(s) F(s, a) - A_\psi(s, a))\partial_\psi A_\psi(s, a).
\end{align}
Both $V$ and $A$ are uniformly lipschitz continuous in $\theta$ and $\psi$, thus
$F$ is lipschitz continuous.

>>>>

%! TEX root = supplementary.tex
Let $(s_t, a_t)_{t\geq 0}$ be the trajectory on which parameters are
learnt. To simplify notations, define
\begin{equation}
	A_\psi(s, a) = \bar{A}_\psi(s, a) - \bar{A}_\psi(s, \pi(s)).
\end{equation}
Define $F$ as
\begin{align}
	F^\theta(\theta, \psi, s, a) &= \alpha^V (r(s, a) + \ln(\gamma) V_\theta(s) + \partial_s V_\theta(s) F(s, a) - A_\psi(s, a))\partial_\theta V_\theta(s)\\
	F^\psi(\theta, \psi, s, a) &= \alpha^A (r(s, a) + \ln(\gamma) V_\theta(s) + \partial_s V_\theta(s) F(s, a) - A_\psi(s, a))\partial_\psi A_\psi(s, a).
\end{align}
From the bounded Hessians and Gradients hypothesis, $V$, $A$, $\partial_s V$,
$\partial_\theta V$ and $\partial_\psi A$ are uniformly Lipschitz continuous in
$\theta$ and $\psi$, thus $F$ is Lipschitz continuous.
